---
title: "BIOL4408 Marine Ecology: Lobster density 1."
author: "TimLanglois"
date: "09/01/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Import lobster density data from a google sheet

First we load some librarys
```{r libs1, message=FALSE, warning=FALSE}
library(googlesheets4) #to read gsheet
library(tidyr) #to tidy data
library(dplyr) #to transform data
library(readr) #to write data
library(here) #to make robust links to files
```

It is useful to set a study name
```{r name, message=FALSE, warning=FALSE}
study<-"lobster.density"
```

Read in the data from the google sheet and check it
```{r read, message=FALSE, warning=FALSE}
#We need the URL for the google sheet
url <- "https://docs.google.com/spreadsheets/d/1Wqn7m2jopx11n5fdl9MHZAujRVkjBusdmIHq_gMhf0A"

# Then we can read it in
dat<-read_sheet(url, sheet = "lobster.density")%>%
  as_tibble()%>%
  select(-c('longitude','latitude','time','way.point','gps','depth','group'))%>%
  glimpse()
```


Now we need to write the data.

There are several ways to setting the directory where we will read data from or write data to.

We are going to use the here() function.

As long as the names on the folders are consistent this function will enable us to work across computers and operating systems.

Let's use here() to make a shortcut to the "Data" directory.
```{r here, message=FALSE, warning=FALSE}
data.dir <- here("Analysis_Lobster_density","Data")
```


Now to write the data we have imported from the googlesheet.
We will append the study name.
```{r write, message=FALSE, warning=FALSE}
setwd(data.dir)#this is our shortcut using here()
dir()

write_csv(dat,paste(study,"gsheet","csv",sep = "."))

```

