---
title: "BIOL4408 Marine Ecology: Lobster density 2."
author: "TimLanglois"
date: "09/01/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 2. Check the data and make corrections.

Load extra librarys
```{r libs2, message=FALSE, warning=FALSE}
library(tidyr) #to tidy data
library(dplyr) #to transform data
library(readr) #to write data
library(here) #to make robust links to files

library(forcats) #to transform catagorical data
library(ggplot2) #to plot data
```


Set a study name
```{r name, message=FALSE, warning=FALSE}
study<-"lobster.density"
```


Use here() to make a shortcut to the "Data" directory.
```{r here, message=FALSE, warning=FALSE}
data.dir <- here("Analysis_Lobster_density","Data")
```


Read in the data and glimpse the data.
glimpse() shows the variable names, formats, head of the data.
```{r read, message=FALSE, warning=FALSE}
setwd(data.dir)#this is out shortcut using here()
dir()

gsheet.dat<-read_csv("lobster.density.gsheet.csv")%>%
  glimpse()

```
This is wide format data.


Check for unique levels of status.
```{r check status, message=FALSE, warning=FALSE}
unique(gsheet.dat$status)
```
there is a typo we need to correct in No-Take


Check for unique levels of sanctuary.
```{r check sanct, message=FALSE, warning=FALSE}
unique(gsheet.dat$sanctuary)

```
We need to make these consistent.


Check for unique levels of site.
```{r check site, message=FALSE, warning=FALSE}
unique(gsheet.dat$site)

```
"eastsalmon" should be "Salmon Bay"


Check for unique levels of year.
```{r check year, message=FALSE, warning=FALSE}
unique(gsheet.dat$year)

```
We think the data from 2017 has errors in it. We should remove it.


Check measures of habitat complexity.
As this is a semi-continous variable we will use summary() to see the range.
```{r check complex, message=FALSE, warning=FALSE}
unique(gsheet.dat$complexity)

summary(gsheet.dat$complexity)

```
The maximum value is supposed to be 4. Looks ok.


Check measures of habitat cover.
```{r check cover, message=FALSE, warning=FALSE}
unique(gsheet.dat$algal.cover)

summary(gsheet.dat$algal.cover)

unique(filter(gsheet.dat,year=="2019")$algal.cover)

```
The maximum value is supposed to be 4.
This does not look consistent. Suggest you only use the most recent years of data.


Check what sites are missing between years.
```{r check sites by years, message=FALSE, warning=FALSE}
table(gsheet.dat$site,gsheet.dat$year)
```
"Armstrong Point","Longreach","Rocky Bay","Stark Bay" - are only sampled once and we should remove them.



Check how we have used site names and that they are unique between levels of status?
```{r check sites, message=FALSE, warning=FALSE}
table(gsheet.dat$site,gsheet.dat$status)
```

## Corrections to the data

We have some corrections to make.
It is always good to have our source data as 'correct' as possible.
However, in this case we will make these corrections in our R import script so that we can keep a record of the changes/corrections we are making to the data.

We can use this list of corrections to go back and check and correct the raw data if we want.

Make corrections and re-format.
```{r correct, message=FALSE, warning=FALSE}

correct.dat<-gsheet.dat %>%
  dplyr::mutate(status = fct_recode(status,
                                    "No-take" = "No-Take"))%>%
  dplyr::mutate(sanctuary = fct_recode(sanctuary,
                                    "Parker Point" = "Parker_Pt",
                                    "Green Island" = "Green_Island",
                                    "Armstrong Bay" = "Armstrong"))%>%
  
#recode the site names
  
    dplyr::mutate(site = fct_recode(site,
                           "Salmon Bay" = "eastsalmon",
                           "City of York" = "City of York Bay",
                           "Ricey Beach" = "Ricey_Bay",
                           "Salmon Bay" = "East_Salmon",
                           "Little Salmon" = "Little_Salmon",
                           "Parker Point" = "Poc_Reef",
                           "Green Island" = "Green_Island",
                           "Parakeet Bay" = "Parakeet_Bay",
                           "Salmon Bay" = "West_Salmon_Bay",
                           "Little Armstrong" = "Little_Armstrong_Bay",
                           "Geordie Bay" = "Geordie_Bay",
                           "Mary Cove" = "Mary_Cove",
                           "Strickland Bay" = "East_Strickland"))%>%
    
  #filter out sites only done once
    filter(!site%in%c(
    "Armstrong Point",
    "Longreach",
    "Rocky Bay",
    "Stark Bay"))%>%
  
  #filter out suspicous year  
    filter(!year==2017)%>%
  
  #remove the levels for the filtered facotrs
    droplevels()%>%
  
  # Make a new unique Site name - to account for Parker Point being used for inside and outside the NTZ.
  dplyr::mutate(site.new=paste(site,status,sep="."))%>%
  
    glimpse()

```


Now we can check our corrections
```{r recheck, message=FALSE, warning=FALSE}

glimpse(correct.dat)

unique(correct.dat$sanctuary) 

unique(correct.dat$site) 

unique(correct.dat$status)

table(correct.dat$site.new,correct.dat$status)


```


Now make new varialbes for sum of legal and sub.legal.
Make the data long format.
```{r make sum}

dat<-correct.dat%>%
  
  replace(is.na(.), 0)%>%
  
  mutate_at(vars(starts_with("x")), funs(ifelse(is.na(.),0,.)))%>%
  mutate_at(vars(starts_with("unsi")), funs(ifelse(is.na(.),0,.)))%>%
  mutate_at(vars(starts_with("legal")), funs(ifelse(is.na(.),0,.)))%>%
  mutate_at(vars(starts_with("sub")), funs(ifelse(is.na(.),0,.)))%>%
  
  # make the legal sum
  dplyr::mutate(legal=(legal.unsized+x80+x85+x90+x95+x100+x105+x110+x115+x120+x125+x130+x135+x140+x145+x150))%>%
  
  # make the sub.legal sum
  dplyr::mutate(sub.legal=(sublegal.unsized+x25+x30+x35+x40+x45+x50+x55+x60+x65+x70+x75))%>%
  
  # make an all sum
  dplyr::mutate(all=(legal+sub.legal+unsized))%>%
  
  # make a unique sample number
  dplyr::mutate(sample.no=1:nrow(.))%>%
  
  # select the variables of interest
  select(c(sample.no,year,date,sanctuary,status,site.new,complexity,algal.cover,legal,sub.legal,all))%>%
  
  # # # make the data long
  gather(key="size.class",value="count",legal, sub.legal,all)%>%
  glimpse()

```

Write the data
```{r message=FALSE, warning=FALSE}
setwd(data.dir) #set the directory
dir() #look in the directory

# Write dat using study name
write_csv(dat,paste(study,"csv",sep = "."))

```


[Go to Lobster denisty 3. Basic plots to check the data](https://github.com/TimLanglois/BIOL4408/blob/c5aa4f8881ad41e1352c8618091e4de530171b0f/Analysis_Lobster_density/3_lobster-density_basic-plots-to-check-data.md)




